Pyg4ometry: a Python library for the creation of Monte Carlo radiation transport physical geometries

### 2020-03-02 public/paper/pyg4ometry.tex
fluka paper parts

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}

@@

\section{FLUKA to Geant4 conversion}
FLUKA geometry is based upon a limited set of primitives (referred to
as \textit{bodies} which can be combined using Boolean operations.  A
\textit{zone} consists of one or more bodies combined using
intersections and subtractions.  Zones may then be further combined
using union operations to form \textit{regions}, which is defined as
the combination of one or more zones as well as a material.  \todo{nested zones.}

Translating FLUKA geometry to GDML first starts with

Translating FLUKA geometry to GDML first starts with a mapping of
FLUKA bodies to GDML solids, which is shown in
Table~\ref{tab:Fluka2Geant4}.

It is notable that FLUKA makes
widespread use of primitives that are infinite in extent, where Geant4

FLUKA geometry makes widespread use of primitives that are infinite in
extent, with this feature distinguishing it most from Geant4 where all
primitives are finite.

FLUKA geometry is most distinuished from Geant4 by its use of
primitives infinite in extent.

Bodies can be combined into
\textit{zones} using intersections and subtractions, with

Bodies may be
combined into \textit{zones} intersections and subtractions, and these
zones may then be combined into \textit{regions} with union
operations.  The most important difference between Geant4 and FLUKA's
implementation of CSG is that FLUKA makes widespread use of primitives
that are infinite in extent, whereas all Geant4 primitives are finite.
Whereas some FLUKA primitives have trivial Geant4 analogues, for
example a sphere, translating the infinite bodies to Geant4 solids is
more involved.  The mapping between FLUKA bodies and Geant4 solids is
shown in Table~\ref{tab:Fluka2Geant4}.

Many FLUKA primitives are infinite in extent, so to emulate
FLUKA geometry construction large but finite extent Geant4 primitive
solids are used.  shows the mapping
between FLUKA and Geant4 objects.  An ANTLR based expression parser is
used to load and comprehend the FLUKA input syntax, an example is
shown in XXX. Each FLUKA body is matched to a Geant4 solid and a FLUKA
region is represented by a Geant4 CSG tree is formed using the
pyg4ometry classes.

@@

\begin{enumerate} 
\item length safety
\item infinite planes
\item disjoint unions 
\end{enumerate}

\section{Geant4 to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 solids can be mapped to a FLUKA CSG region.
Some solids in Geant4 directly map to a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between geant4 solids and the bodies used to compose a fluka region.

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{siunitx}

@@

\section{FLUKA to GDML conversion}
FLUKA geometry is based upon a limited set of primitives (referred to
as \textit{bodies} which can be combined using Boolean operations.  A
\textit{zone} consists of one or more bodies or \textit{subzones}
combined using intersections and subtractions.  Zones may then be
further combined using union operations to form \textit{regions},
which is defined as the union of one or more zones, as well as a
material.

Each FLUKA body is represented in Pyg4ometry with a corresponding
class, and in turn each class has methods which returns a GDML
primitive solid and that solid's rotation and position such that it
matches its FLUKA equivalent.  The expansion, translation and
transform geometry directives are each folded into one or more of
these three methods.  The mapping of FLUKA bodies to GDML solids is
shown in Table~\ref{tab:Fluka2Geant4}.  It is worth noting that many
of the FLUKA bodies are infinite in extent, but are mapped to finite
GDML solids.  The translation of infinite bodies to equivalent finite
solids is one of the main steps involved in the conversion process.
This mapping is possible because whilst FLUKA bodies can be infinite
in extent, all zones and regions must be finite.  Zones and regions
are then composed by instantiating these classes and adding body
instances to them.  Each Zone and Region instance can then return its
equivalent GDML Boolean solid.

The FLUKA CSG ASCII format is parsed using an ANTLR4-generated parser,
producing an abstract syntax tree (AST).  The resulting AST is then
walked to populate Region instances with zones and bodies.  An example
FLUKA region with the corresponding abstract syntax tree is shown in
Figure~\ref{fig:region-ast}.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=7cm]{./diagrams/rapidModelling.pdf}
\caption{FLUKA input with an AST.}
\label{fig:region-ast}
\end{center}
\end{figure}

@@

\subsection{Infinite bodies}

The majority of bodies in FLUKA are infinite in extent, and fall
broadly into four categories, half-spaces, infinitely long cylinders,
infinitely long elliptical cylinders and quadric surfaces.
Translating these bodies to Geant4 requires generating the equivalent
finite solid whilst retaining the same final finite Boolean shape.
This is achieved with the use of axis-aligned bounding boxes (AABB),
in which the FLUKA body is translated to a finite solid with with
dimensions slightly larger than the AABB.  The lengths of infinite
(elliptical) cylinders are reduced to finite equivalents with lengths
slightly greater than the bounding box.  Similarly, half spaces are
reduced to boxes with one face acting as that of the half-space face,
and quadric surfaces are sampled only over the volume denoted by the
AABB.  Furthermore, the positions of these solids are moved as close
to the bounding box as possible whilst retaining an identical final
Boolean geometry.

Generating these bounding boxes over which the bodies should be
translated involves first evaluating each region with very large Tubs,
EllipticalTubes and Boxes (by default \SI{50}{\km} in length), such
that they are effectively infinite for most reasonable use cases.
Pyg4ometry's CSG meshing is then used to generate a mesh for each
region, from which the axis-aligned bounding box can be extracted.
Each region is then evaluated a second time with respect to its
respective bounding box, with all of its constituent infinite solids
being reduced as described above.  This algorithm works robustly for
infinite (elliptical) cylinders and half spaces as the number of
facets is independent of the size of the solid, so generating the
initial mesh from large solids works well.  Generating a quadric
surface over a very large volume in space whilst retaining topological
information is computationally very expensive, so to resolve this the
user must provide the approximate axis-aligned bounding box of any
region in which a quadric is used.

\subsection{Coplanar faces}

Coplanar faces, in which the faces of two union components or that of
two regions are perfectly coplanar in FLUKA are ubiquitous and present
no difficulties to the operation of the program.  However, in Geant4
these will generally result in tracking errors.  These must be handled
robustly to ensure the resulting geometry is usable.  Coplanar faces
are resolved automatically by slightly decreasing the size of every
body that is used in an intersection, and increasing the size of every
body used in a subtraction.  These rules are inverted for nested
subtractions and work well for guaranteeing well-formed geometry free
from overlaps.

\subsection{Removing redundant half-spaces}

\subsection{Disjoint unions}

\subsection{Lattice}

\subsection{Conclusion?}

\begin{itemize}
\item \item disjoint subtractions?
\item quadric specialisation
\item second order lattice
\end{itemize}
\section{Geant4 to FLUKA conversion}
It is relatively straight forward to convert Geant4 geometry to FLUKA. Each of the Geant4 solids can be mapped to a FLUKA CSG region.
Some solids in Geant4 directly map to a FLUKA body, others need a construction of a simple FLUKA CSG tree. Table~\ref{tab:geant2fluka}
lists the mapping between geant4 solids and the bodies used to compose a fluka region.
